Ear canal dynamic motion piezoelectric energy harvester using a bistable resonator cycled by coupled hydraulic switches made of collapsed flexible tubes.

### 2022-10-10 article_system.tex
update

\title{Ear canal dynamic motion piezoelectric energy harvester using a bistable oscillator cycled by coupled hydraulic switches made of collapsed flexible tubes.}

@@

The common energy sources are solar or wind energy but their availability in the low size environment of the application context makes them not suitable for exploitation. The activities of the human body generate a considerable amount of energy of different natures. Among them, the electrochemical energy from the inner ear \cite{Mercier2012}, the kinetic energy from walking or head movements \cite{Azimi2021,Smilek2016}, the body heat thermal energy \cite{Kim2014}, the strain energy from the skin deformation \cite{Jin2021} could be harvested for in-ear applications. An interesting energy source for cochlear implants or hearing aids appears to be the ear canal mechanical deformation as it is directly located in the area of the application. Delnavaz \emph{et al.} first showed in 2012 that the jaw movements changes de ear canal geometry during mastication. Carioli \emph{et al.} then used a personalized earplug molding technic with a 3D scanner to reveal a global bending movement and a local compression area in the ear canal during mastication \cite{Carioli2016} (fig. \ref{fig:molding}). These two mechanical deformations represent an available energy source that can be exploited. Besides this work, the literature exposes two technological solutions adapted to the extraction of the mechanical deformation energy in the ear canal.
\begin{figure*}[!htbp]
	\includegraphics[trim={6cm 0cm 0cm 10cm},clip, width=0.6\linewidth]{figures/molding.pdf}
	\caption{(a) Earcanal in two extreme positions showing a global bending movement of the earcanal (b) Earcanal sectional view in two extreme positions showing a local compression area of the earcanal. \cite{Carioli2016}}
	\label{fig:molding}
\end{figure*}
Delnavaz \emph{et al.} first developed a flexible piezoelectric harvester made of a silicon earplug with a PVDF layer \cite{Delnavaz2013} (fig. \ref{fig:critias_piezo}). The earplug has been melted on the subject in order to maximize the contact surface with the ear canal. The harvester is capable of exploiting both the flexion and the compression energy sources. The experimental prototype presented on figure \ref{fig:critias_piezo} was able to extract $44$µJ from a normal mastication cycle. The soft materials facilitate the bio-integration of the harvester but the PVDF low electromechanical coupling coefficient could have led to poor energy conversion efficiency. A hydro-electromagnetic energy harvester was also developed by the same researchers \cite{Delnavaz2012}. The energy is extracted by a liquid filled earplug whose internal volume and pressure vary as a result of the radial compression of the ear canal internal walls during the jaw movements. These variations put into motion a moving magnet translating in a water column surrounded by a copper coil. The energy was harvested by electromagnetic induction and experimentally evaluated at 0.2µJ per mastication cycle. The prototype appeared to be less efficient than the PVDF earplug. This could be explained by the poor power generation of the electromagnetic transduction for a low frequency energy source (1.57Hz) and a small scale application \cite{Kulah2008,Priya2017}. However, the hydraulic transmission interface allows the use of more complex and voluminous technological solutions since the energy becomes available outside the small ear canal volume. The table \ref{tab:harvesters_ear} summarizes the existing harvesters that exploit the ear canal mechanical deformation energy. \\
	\includegraphics[trim={18cm 0cm 0cm 8.7cm},clip, width=\linewidth]{figures/critias_piezo.pdf}
	\caption{Piezoelectric energy harvester. Experimental setup (dimensions are in mm). (a) Flattened PVDF layer. (b) PVDF layer. (c) Prototype of piezo-earpiece. (d) Piezo-earpiece mounted on the headset. (e) Measuring setup. \cite{Delnavaz2013}}  
	\label{fig:critias_piezo}
	\includegraphics[trim={18cm 0cm 0cm 8.75cm},clip, width=\linewidth]{figures/critias_emag.pdf}
	\caption{Hydroelectromagnetic energy harvester. (a) Test setup. (b) Energy harvesting module. \cite{Delnavaz2012}} 
	\label{fig:critias_emag}
	\begin{tabular}{ l  c  c  c}
		\multicolumn{1}{l}{\textbf{Harvester}}                        &
		\multicolumn{1}{c}{\textbf{Harvested energy per mastication}} &
		\multicolumn{1}{c}{\textbf{Reference}}                        \\
		Piezo earpiece        & $44$µJ 		 & \cite{Delnavaz2013} \\
	    Hydroelectromagnetic  & $0.2$µJ      & \cite{Delnavaz2012} \\
		Piezo sensor          & $0.1$µJ      & \cite{Carioli2018}  \\
	\caption{Existing harvesters exploiting the ear canal deformation energy}
	\label{tab:harvesters_ear}
Based on the literature, we propose a new technological solution optimizing the energy conversion efficiency to maximize the harvested energy from the ear canal dynamic motion. The next section states on the energy source characterization and on the harvesting strategy we adopted. Section \ref{sec:HARVESTER PRESENTATION AND OPERATION PRINCIPLE} presents the harvester and describes its operation principle. Section \ref{sec:SYSTEM MODELING AND SIMULATIONS} shows the global system  modeling and the numerical simulation on its behavior. Then section \ref{sec:EXPERIMENTAL CHARACTERIZATIONS} then exposes the experimental validation performed on the harvester critical components and section \ref{sec:MODEL RECALIBRATION WITH EXPERIMENTAL DATA} presents the experimentally adjusted global harvester model. Section \ref{sec:ANALYZE AND DISCUSSION} discusses this work major contributions and the harvester potential. Finally, a conclusion is stated in the last section.

%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
\section{ENERGY SOURCE AND HARVESTING STRATEGY}
\label{sec:THE ENERGY SOURCE AND HARVESTING STRATEGY}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
    %///////////////////////////////////////////// 
	\subsection{The energy source}	
	\label{The energy source}
    %/////////////////////////////////////////////
The ear canal deflection energy is not a usual source because of the soft tissues and the low frequency of mastication. Thus, the traditional transducing methods are not suitable enough for the application.

On one hand, the ear canal global bending energy is 5 times higher than the local compression energy. To date, the literature proposes only one way to harvest the bending energy from the ear canal wall, which is to use of soft materials as the piezoelectric earplug previously introduced \cite{Delnavaz2013}. The very low mechanical impedance of the soft tissue in fact limits the use of more effective rigid piezoelectric materials.

On the other hand, the hydraulic extraction strategy proposed in \cite{Delnavaz2012} only exploits the local compression energy. However, it allows to transmit the energy outside the ear canal and benefit so from larger effective space. The latter is in fact essential as it becomes then possible to implement an impedance adaption stage and more effective transducing methods that could not fit directly inside the ear canal restricted volume. The present work is then based on the exploitation of the ear canal by a liquid filled earplug. As energy source, the latter can be assimilated to a micro pump characterized by its volume variation $V_{ear}$ and pressure variation $p_{ear}$. In order to optimize the energy extraction, the earplug must fit the ear canal wall. Turcot \emph{et al.} demonstrated that the initial pressurization must be at a minimum of 14kPa to achieve a good fit. Bouchard-Roy \emph{et al.} studied the dynamic pressure variation inside the earplug for different initial pressurization levels \cite{Bouchard-Roy2020}. A maximum of $\text{max}(\Delta p_{ear})=12$kPa of dynamic pressure variation , with a $28$kPa of initial pressurization, has been noted during the mastication for seven subjects. Besides, the temporal evolution of the volume variation $\Delta V_{ear}$ has been estimated in \cite{Delnavaz2012} and the result for one mastication cycle is presented on figure \ref{fig:deltaV_ear}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={20.5cm 0cm 0cm 10.8cm},clip, width=0.4\textwidth]{figures/deltaV_ear.pdf}
	\caption{Ear canal volume variation for one mastication cycle \cite{Delnavaz2012}}
	\label{fig:deltaV_ear}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The earplug operation characteristic behavior has not been studied yet. Thus, the evolution of the volume variation $\Delta V_{ear}$ is not well known as a function of the pressure variation $\Delta p_{ear}$. However, considering the pressure and volume variation levels from \cite{Delnavaz2012} and \cite{Bouchard-Roy2020}, the equation \ref{eq:max_hydraulic_energy} approximates an upper limit for the extractable hydraulic energy with a 28kPa pressurized earplug and a maximum volume variation of $\Delta V_{ear,m}=60$mm$^3$. $p_c$ stands for the comfort pressure by estimating that the user could feel discomfort if $\Delta p_{ear}>p_c$. The energy estimation assumes that the earplug is a perfect flow rate source, which is an optimistic hypothesis leading to $0.9$mJ available from one mastication. Based on 2200 mastication cycles per day, according to \cite{Goll2011}, there would be $2$mJ available per day, per ear. This amount of energy could theoretically enhance supply 23.5\% of the 17J of needed for the state of the art of cochlear implants \cite{Kulah2022}. Further research in both domains related to hearing aids and energy harvesting could lead to fully autonomous devices in the future.
\begin{align}
	\text{max}(E_{hyd}) = p_c \cdot \Delta V_{ear,m} = 0.9\text{mJ}
	\label{eq:max_hydraulic_energy}
\end{align}

    %///////////////////////////////////////////// 
	\subsection{The harvesting strategy}	
	\label{The harvesting strategy}
    %/////////////////////////////////////////////
The application is not suitable for magnetic interactions for two major reasons. Magnetic fields admit high non-linear properties at small scales and their effectiveness decreases as the effective volume of transducing material and the frequency of the energy source decrease \cite{Priya2017,Kulah2008}. The piezoelectric ceramics yet appear like good candidates, but they can not be easily used because of their high stiffness. For this reason, and also because of the higher conversion efficiency of resonant structures, the harvesting system could benefit from a frequency-up conversion stage \cite{Ashraf2011,Peng2021}. These structures are commonly used for broadband energy harvesting. A majority of the papers with piezoelectric transducers use mechanical stops with piezoelectric ceramic cantilevers to reach high frequencies with a good mechanical coupling \cite{Edwards2013,Gu2011,Lee2007}. Stoppers however dissipate energy and demand maintenance with time, in order to replace rubbing parts. To overcome this drawback, alternative solutions use mechanically bistable oscillators (BO) to perform frequency-up without any rubbing parts \cite{Vocca2012}. Recent advances 
led to the development and optimization of  bistable structures integrating stacked piezoelectric ceramics inside flextensional elastic structures \cite{Huguet2017}. They show enhanced performances due to high electromechanical coupling and broadband efficiency.
Preliminary calculations based on the Duffing equation defining the BO mechanical behavior \cite{Tseng1971}, and the linearity of piezoelectric beams, offer a comparison tool to determine the best technological approach to maximize the harvested energy from the hydraulic source. Figure \ref{fig:OB_vs_LINEAR} shows in blue the hydraulic behavior of the ideal flow rate source previously introduced. To be harvested, the energy has to be stored in a mechanical potential reservoir at first, to be then converted in electricity by the transducer. The hydraulic-mechanic interface is chosen to be a hydraulic cylinder. On one hand, the best linear spring is capable of storing/extracting a maximum of $0.5\%$ from the maximum available hydraulic energy ($\text{max}(E_{hyd})$) if actuated by the source through a hydraulic cylinder. On the other hand, a BO type non-linear spring is inherently capable of extracting more than $65\%$ from the same source.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={22.5cm 0cm 0cm 8.8cm},clip, width=0.35\textwidth]{figures/OB_vs_LINEAR.pdf}
	\caption{Hydraulic source exploitation depending on the mechanical load}
	\label{fig:OB_vs_LINEAR}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Hence, the absence of rubbing parts and the optimal energy extraction shed light on a BO as a promising solution for our application. 
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
\section{HARVESTER PRESENTATION AND OPERATION \mbox{PRINCIPLE}}
\label{sec:HARVESTER PRESENTATION AND OPERATION PRINCIPLE}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={3.2cm 0cm 0cm 4.3cm},clip, width=0.7\textwidth]{figures/system_presentation.pdf}
	\caption{Schematic presentation of the frequency-up piezoelectric harvester exploiting the ear canal geometry variation} 
	\label{fig:system_presentation}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The figure \ref{fig:system_presentation} presents a schematic view of the energy harvesting system. It uses a liquid filled earplug to transmit the energy outside the ear canal and benefit so from larger effective space. To maximize the efficiency, the system integrates a frequency-up conversion stage performed by a bistable oscillator (BO). The latter implements an amplified piezoelectric generator (APG) made of a flextensional elastic APX4 steel structure containing stacked lead zirconate titanate (PZT) ceramics (fig. \ref{fig:APG}). The electromechanical converter composed of the BO and the APG theoretically offers higher energy conversion efficiency compared to low frequency solutions and/or flexible electromechanical transducers \cite{Guo2019}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
\begin{figure}[!htb]
	\begin{center}
		\begin{subfigure}[t]{0.6\linewidth}
			\captionsetup{justification=centering}
			\includegraphics[trim={16cm 0cm 0cm 6cm},clip,width=\linewidth]{figures/APG_schema.pdf}
			\caption{Detailed schema}
			\label{fig:APG_schema}
		\end{subfigure}
		\hfillx
		\begin{subfigure}[t]{0.29\linewidth}
			\captionsetup{justification=centering}
			\includegraphics[trim={29.5cm 0cm 0cm 8cm},clip,width=0.65\linewidth]{figures/APG_photo.pdf}
			\caption{Picture}
			\label{fig:APG_photo}
		\end{subfigure}
		\caption{The APG presentation}
		\label{fig:APG}
	\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

To maximize the mechanical energy transfer between the ear canal and the earplug, the latter is lightly pressurized. A minimum of 14kPa is needed in order to fit the earplug to the ear canal wall \cite{Delnavaz2013}. The pressure separator ensures that this initial pressurization does not interfere on the system operation. Then, during a mastication cycle, the TMJ compresses the ear canal wall and expels so the fluid outside the earplug. The HVs ensure to led the fluid alternatively to each HC in order to actuate the BO mass back and forward. The hydraulic amplifier adapts the cylinders stroke to the available earplug flow rate and also adapts the mechanical impedance of the BO to the mechanical impedance of the ear canal wall.

The BO mass has two stable equilibrium positions, $x_m = x_0$ and $x_m = -x_0$. The harvester operates on two major phases. The first consists on the BO mass actuation, by a HC, from one stable equilibrium position toward the unstable position at $x_m = 0$. During the second phase the mass goes oscillate around the opposite stable equilibrium position until it stops, while a portion of the oscillatory energy is converted in electricity by the GPA. The mass then moves backward during the next mastication, completing so the harvester cycle.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={20cm 0cm 0cm 10.75cm},clip, width=\linewidth]{figures/HV_actuation_detail.pdf}
	\caption{Contact detail between a HV and the BO mass} 
	\label{fig:HV_actuation_detail}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In order to maximize the harvested energy we must minimize the energy consumed by the HVs. A technological solution for this role is to use a flexible tube buckled by bending. The figure \ref{fig:HV_actuation_detail} schematizes the integration of such HV around the BO environment. The valve is composed of a flexible tube connected between a rigid part of the hydraulic circuit and a fixed one. The BO mass motion on the mobile part induces a bending angle $\theta$ around the rotational point $O$ located at the buckled section of the flexible tube. Consequently, a pressure loss is generated through the valve in buckled position. Thus, a HV is considered closed when the BO mass is at a stable equilibrium position ($x_m=\pm x_0$) and opened when the mass is at $x_m=0$. Two HVs symmetrically placed on each side of the $\vec{x}$ axis set up an opened hydraulic circuit on one side and a closed one on the other side, depending on the BO mass position. The HV technological choice is motivated by three major arguments. First, the absence of electromechanical transduction minimizes the energy losses for the valve operation. Then, the post-buckling softening of the bent tube minimizes the mechanical impact on the BO mass dynamic operation. Finally, it is well adapted at the millimetric application scale.

%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
\section{SYSTEM GLOBAL MODELING}
\label{sec:SYSTEM MODELING}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
The harvester is a multyphisic system composed of hydraulic, mechanical, and electrical components. This section presents a global multiphysic model describing the theoretical system operation. Two coupled subsystems are considered. The subsection \ref{subsec:The electromechanical converter} shows the modeling of the electromechanical frequency-up converter (BO + APG) under the mechanical influence of a HV and a HC. Then, the subsection \ref{subsec:The hydraulic circuit} discusses the design of the hydraulic circuit composed of the two HVs, the two HCs, the hydraulic amplifier, the pressure separator and the liquid filled earplug. Finally, the section \ref{subsec:Global model simulation} exposes the numerical simulation of the designed multiphysic coupled model.

@@

The simulated model shows promising efficiency and demonstrates that a minimum value of 26 is required on the hydraulic restriction coefficient $r_{Cf}$ to ensure the adequate cycled actuation of the BO mass. Next section will present the experimental approach to validate the theoretical behavior of the electromechanical converter.

>>>>

\title{Ear canal dynamic motion piezoelectric energy harvester using a bistable resonator cycled by coupled hydraulic switches made of collapsed flexible tubes.}

@@

The common energy sources are solar or wind energy but their availability in the low size environment of the application context makes them not suitable for exploitation. The activities of the human body generate a considerable amount of energy of different natures. Among them, the electrochemical energy from the inner ear \cite{Mercier2012}, the kinetic energy from walking or head movements \cite{Azimi2021,Smilek2016}, the body heat thermal energy \cite{Kim2014}, the strain energy from the skin deformation \cite{Jin2021} could be harvested for in-ear applications. An interesting energy source for cochlear implants or hearing aids appears to be the ear canal mechanical deformation as it is directly located in the area of the application. Delnavaz \emph{et al.} first showed in 2012 that the jaw movements changes de ear canal geometry during mastication. Carioli \emph{et al.} then used a customized earplug molding technic with a 3D scanner to reveal a global bending movement and a local compression area in the ear canal during mastication \cite{Carioli2016}. These two mechanical deformations represent a potential energy source that can be exploited.

Delnavaz \emph{et al.} developed a flexible piezoelectric harvester made of a silicon earplug with a PVDF layer \cite{Delnavaz2013}. The earplug has been molded on the subject in order to maximize the contact surface with the ear canal. The harvester is capable of exploiting both the flexion and the compression energy sources. The experimental prototype was able to extract $44$µJ from a mastication cycle. The soft materials facilitate the bio-integration of the harvester but the PVDF low electromechanical coupling coefficient could have led to poor energy conversion efficiency. A hydro-electromagnetic energy harvester was also developed by the same researchers \cite{Delnavaz2012}. The energy is extracted by a liquid filled earplug whose internal volume vary as a result of the radial compression of the ear canal internal walls during the jaw movements. This variation put into motion a moving magnet translating in a water column surrounded by a coil. The generated energy experimentally amounts to 0.2µJ per mastication cycle. The prototype appeared to be less efficient than the PVDF earplug. This could be explained by the poor power generation of the electromagnetic transduction for a low frequency energy source (1.57Hz) and a small scale application \cite{Kulah2008,Priya2017}. However, the hydraulic transmission allows the use of more complex and voluminous technological solutions since the energy becomes available outside the small ear canal volume.

Based on the analysis of the literature, we propose a new solution optimizing the energy conversion efficiency to maximize the harvested energy from the ear canal dynamic motion. The next section states on the hydraulic energy source characterization and on the harvesting strategy we adopted. Section \ref{sec:HARVESTER PRESENTATION AND OPERATION PRINCIPLE} presents the harvester and describes its operation principle. Section \ref{sec:SYSTEM MODELING AND SIMULATIONS} shows the global system  modeling and the numerical simulation on its behavior. Then section \ref{sec:EXPERIMENTAL CHARACTERIZATIONS} then exposes the experimental validation performed on the harvester critical components and section \ref{sec:MODEL RECALIBRATION WITH EXPERIMENTAL DATA} presents the experimentally adjusted global harvester model. Section \ref{sec:ANALYZE AND DISCUSSION} discusses this work major contributions and the harvester potential. Finally, a conclusion is stated in the last section.

%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
\section{ENERGY SOURCE AND HARVESTING STRATEGY}
\label{sec:THE ENERGY SOURCE AND HARVESTING STRATEGY}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
    %///////////////////////////////////////////// 
	\subsection{The energy source}	
	\label{The energy source}
    %/////////////////////////////////////////////
Human tissue requires soft flexible materials in order to extract the ear canal deflection energy without causing discomfort. However, flexible transducers admit low energy conversion efficiency as a part of the extracted energy is dissipated in the elastic non-active material. Additionally, the low frequency of mastication makes it difficult to use resonant structures that generally admit better conversion efficiencies than low frequency solutions \cite{Ashraf2011}. Thus, the traditional transducing methods are not suitable enough for the application.

To date, the best solution proposed in the literature consists on using soft materials as the piezoelectric PVDF earplug previously introduced \cite{Delnavaz2013}. The low mechanical impedance of the soft tissue in fact limits the use of more effective rigid piezoelectric materials.

The hydraulic extraction strategy proposed in \cite{Delnavaz2012} is less effective than the piezoelectric earplug but allows to transmit the energy outside the ear canal and benefit so from larger effective space. The latter is in fact essential as it becomes then possible to implement an impedance adaption stage and more effective transducing methods that could not fit directly inside the ear canal restricted volume. In fact, Marin \emph{et al.} demonstrated that the power generated by piezoelectric or electromagnetic transducers is as much better as the effective volume of transducing material is high \cite{Marin2011}. The present work is then based on the exploitation of the ear canal by a liquid filled earplug. As energy source, it can be assimilated to a micro pump characterized by its volume variation $V_{ear}$ and pressure variation $p_{ear}$. In order to optimize the energy extraction, the earplug must fit the ear canal wall. Turcot \emph{et al.} demonstrated that the initial pressurization must reach a minimum of $14$kPa to achieve a good fit \cite{TURCOT2011}. Bouchard-Roy \emph{et al.} studied the dynamic pressure variation inside the earplug for different initial pressurization levels \cite{Bouchard-Roy2020}. A maximum of $\text{max}(\Delta p_{ear})=12$kPa of dynamic pressure variation, with a $28$kPa of initial pressurization, has been noted during the mastication for seven subjects. Besides, the temporal evolution of the volume variation $\Delta V_{ear}$ has been estimated in \cite{Delnavaz2012} and the result for a mastication cycle is presented on figure \ref{fig:deltaV_ear}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={20.5cm 0cm 0cm 10.8cm},clip, width=0.4\textwidth]{figures/deltaV_ear.pdf}
	\caption{Ear canal volume variation for one mastication cycle}
	\label{fig:deltaV_ear}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The earplug operation characteristic behavior has not been studied yet. Thus, the evolution of the volume variation $\Delta V_{ear}$ is not well known as a function of the pressure variation $\Delta p_{ear}$. Then, considering the pressure and volume variation levels from \cite{Delnavaz2012} and \cite{Bouchard-Roy2020}, the equation \ref{eq:max_hydraulic_energy} approximates an upper limit for the extractable hydraulic energy with the maximum values of $\Delta p_{ear}$ et $\Delta V_{ear}$ that have been recorded, i.e. $p_c=12kPa$ and $\Delta V_{ear,m}=60$mm$^3$. $p_c$ stands for the comfort pressure by assuming that the user could feel discomfort if $\Delta p_{ear}>p_c$. The energy estimation presumes that the earplug is a perfect flow rate source, which is a theoretical hypothesis leading to $0.9$mJ available from one mastication. Based on 2200 mastication cycles per day, according to \cite{Goll2011}, there would be $2$mJ available per day, per ear. This amount of energy could theoretically enhance supply $23.5$\% of the $17$J of needed for the state of the art of cochlear implants \cite{Kulah2022}. Further research in both domains related to hearing aids and energy harvesting could lead to fully autonomous devices in the future.
\begin{align}
	\text{max}(E_{hyd}) = p_c \cdot \Delta V_{ear,m} = 0.9\text{mJ}
	\label{eq:max_hydraulic_energy}
\end{align}

    %///////////////////////////////////////////// 
	\subsection{The harvesting strategy}	
	\label{The harvesting strategy}
    %/////////////////////////////////////////////
Electromagnetic transducers are usually used to harvest energy from kinetic (vibration) energy sources. In fact, the power density of electromagnetic transducers decreases as much as the effective volume of transducing material and the frequency of the energy source decrease \cite{Priya2017,Kulah2008}. Also, the fabrication and the integration of the coil and the moving magnet can come out to be difficult as the gap between the two components has to be carefully managed to prevent non-linear unpredicted responses \cite{Caruntu2001}. Thus, piezoelectric ceramics appear like good candidates, but they can not be easily used because of their high stiffness and their high resonance frequencies (>$1.57$Hz). Therefore, the harvesting system could benefit from a frequency-up conversion stage \cite{Ashraf2011,Peng2021} to increase the frequency of the energy source and so the energy conversion efficiency of piezoelectric ceramic transducers. A majority of the frequency-up systems with piezoelectric transducers use mechanical stops with piezoelectric ceramic cantilevers to reach high frequencies with a good mechanical coupling \cite{Edwards2013,Gu2011,Lee2007}. Stoppers however dissipate energy. Alternative solutions use mechanically bistable resonators (BR) to perform frequency-up \cite{Vocca2012}. Recent works led to the development and optimization of  bistable structures integrating stacked piezoelectric ceramics inside flextensional elastic structures \cite{Huguet2017}. They show enhanced performances due to high electromechanical coupling.

We can estimate the best choice maximizing the energy extraction from the hydraulic source. Figure \ref{fig:OB_vs_LINEAR} shows in blue the hydraulic behavior of the ideal flow rate source previously introduced with the upper limit values of pressure and volume variations. To be harvested, the energy has to be stored in a mechanical potential reservoir at first, to be then converted in electricity by the transducer. The hydraulic-mechanic interface can be a hydraulic cylinder. In this way, the best linear resonator can theoretically exploit/store a maximum of $50\%$ from the maximum available hydraulic energy ($\text{max}(E_{hyd})$). Furthermore, BR usually admit non-linear elastic behavior. For example, a BR governed by a Duffing equation is inherently capable of exploiting $65$\% from the same energy source, which is $15$\% better than the best theoretical linear option.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={22.5cm 0cm 0cm 8.8cm},clip, width=0.35\textwidth]{figures/OB_vs_LINEAR.pdf}
	\caption{Hydraulic source exploitation depending on the mechanical load}
	\label{fig:OB_vs_LINEAR}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Hence, the absence of rubbing parts and the optimal energy extraction shed light on a BR as a promising solution for our application. 
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
\section{HARVESTER PRESENTATION AND OPERATION \mbox{PRINCIPLE}}
\label{sec:HARVESTER PRESENTATION AND OPERATION PRINCIPLE}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={3.2cm 0cm 0cm 4.3cm},clip, width=0.7\textwidth]{figures/system_presentation.pdf}
	\caption{Schematic presentation of the frequency-up piezoelectric harvester exploiting the ear canal geometry variation} 
	\label{fig:system_presentation}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The figure \ref{fig:system_presentation} presents a schematic view of the energy harvesting system. It uses a liquid filled earplug to transmit the energy outside the ear canal and benefit so from larger effective space. To maximize the efficiency, the system integrates a frequency-up conversion stage performed by a bistable resonator (BR). The latter implements an amplified piezoelectric generator (APG) made of a flextensional elastic APX4 steel structure containing stacked lead zirconate titanate (PZT) ceramics (fig. \ref{fig:APG}). The electromechanical converter composed of the BR and the APG theoretically offers higher energy conversion efficiency compared to low frequency solutions and/or flexible electromechanical transducers \cite{Guo2019}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
\begin{figure}[!htb]
	\begin{center}
		\begin{subfigure}[t]{0.6\linewidth}
			\captionsetup{justification=centering}
			\includegraphics[trim={16cm 0cm 0cm 6cm},clip,width=\linewidth]{figures/APG_schema.pdf}
			\caption{Detailed schema}
			\label{fig:APG_schema}
		\end{subfigure}
		\hfillx
		\begin{subfigure}[t]{0.29\linewidth}
			\captionsetup{justification=centering}
			\includegraphics[trim={29.5cm 0cm 0cm 8cm},clip,width=0.65\linewidth]{figures/APG_photo.pdf}
			\caption{Picture}
			\label{fig:APG_photo}
		\end{subfigure}
		\caption{The APG presentation}
		\label{fig:APG}
	\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

To maximize the mechanical energy transfer between the ear canal and the earplug, the latter is lightly pressurized. A minimum of 14kPa is needed in order to fit the earplug to the ear canal wall \cite{Delnavaz2013}. The pressure separator ensures that this initial pressurization does not interfere on the system operation. Then, during a mastication cycle, the TMJ compresses the ear canal wall and expels so the fluid outside the earplug. The HVs ensure to led the fluid alternatively to each HC in order to actuate the BR mass back and forward. The hydraulic amplifier adapts the cylinders stroke to the available earplug flow rate and also adapts the mechanical impedance of the BR to the mechanical impedance of the ear canal wall.

The BR mass has two stable equilibrium positions, $x_m = x_0$ and $x_m = -x_0$. The harvester operates on two major phases. The first consists on the BR mass actuation, by a HC, from one stable equilibrium position toward the unstable position at $x_m = 0$. During the second phase the mass goes oscillate around the opposite stable equilibrium position until it stops, while a portion of the resonatory energy is converted in electricity by the GPA. The mass then moves backward during the next mastication, completing so the harvester cycle.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={20cm 0cm 0cm 10.75cm},clip, width=\linewidth]{figures/HV_actuation_detail.pdf}
	\caption{Contact detail between a HV and the BR mass} 
	\label{fig:HV_actuation_detail}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In order to maximize the harvested energy we must minimize the energy consumed by the HVs. A technological solution for this role is to use a flexible tube buckled by bending. The figure \ref{fig:HV_actuation_detail} schematizes the integration of such HV around the BR environment. The valve is composed of a flexible tube connected between a rigid part of the hydraulic circuit and a fixed one. The BR mass motion on the mobile part induces a bending angle $\theta$ around the rotational point $O$ located at the buckled section of the flexible tube. Consequently, a pressure loss is generated through the valve in buckled position. Thus, a HV is considered closed when the BR mass is at a stable equilibrium position ($x_m=\pm x_0$) and opened when the mass is at $x_m=0$. Two HVs symmetrically placed on each side of the $\vec{x}$ axis set up an opened hydraulic circuit on one side and a closed one on the other side, depending on the BR mass position. The HV technological choice is motivated by three major arguments. First, the absence of electromechanical transduction minimizes the energy losses for the valve operation. Then, the post-buckling softening of the bent tube minimizes the mechanical impact on the BR mass dynamic operation. Finally, it is well adapted at the millimetric application scale.

%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
\section{SYSTEM GLOBAL MODELING}
\label{sec:SYSTEM MODELING}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
The harvester is a multyphisic system composed of hydraulic, mechanical, and electrical components. This section presents a global multiphysic model describing the theoretical system operation. Two coupled subsystems are considered. The subsection \ref{subsec:The electromechanical converter} shows the modeling of the electromechanical frequency-up converter (BR + APG) under the mechanical influence of a HV and a HC. Then, the subsection \ref{subsec:The hydraulic circuit} discusses the design of the hydraulic circuit composed of the two HVs, the two HCs, the hydraulic amplifier, the pressure separator and the liquid filled earplug. Finally, the section \ref{subsec:Global model simulation} exposes the numerical simulation of the designed multiphysic coupled model.

@@

The simulated model shows promising efficiency and demonstrates that a minimum value of 26 is required on the hydraulic restriction coefficient $r_{Cf}$ to ensure the adequate cycled actuation of the BR mass. Next section will present the experimental approach to validate the theoretical behavior of the electromechanical converter.
